---
title: "INTOXICATE Poster Figures"
author: "Yash Jaggi"
date: "WCM EM Research Symposium · May 7, 2026"
output:
  pdf_document:
    latex_engine: xelatex
    toc: false
geometry: "left=0.7in,right=0.7in,top=0.7in,bottom=0.7in"
header-includes:
  - \usepackage{float}
  - \usepackage{caption}
  - \captionsetup[figure]{labelformat=empty,font=normalsize,skip=4pt}
  - \renewcommand{\familydefault}{\sfdefault}
  - \pagestyle{empty}
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
# ============================================================================
# INTOXICATE POSTER FIGURES
# WCM Emergency Medicine Research Symposium · May 7, 2026
#
# Two figures, computed live from the toxicology registry. Each is saved
# as a 300 DPI PNG to ./poster_figures/ for direct insertion into the
# PowerPoint poster template.
# ============================================================================

library(dplyr)
library(tidyr)
library(ggplot2)
library(readxl)
library(scales)
library(stringr)
library(knitr)

knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  fig.align = "center", dpi = 300
)

# ---- WCM PALETTE ----------------------------------------------------------
WCM_MAROON <- "#A8132B"
WCM_NAVY   <- "#1B365D"
INK        <- "#1A1A1A"
SLATE      <- "#4A4A4A"
MIST       <- "#C9C9C9"
PAPER      <- "#FFFFFF"
GRID       <- "#E8E8E8"

# ---- OUTPUT ---------------------------------------------------------------
fig_dir <- "poster_figures"
if (!dir.exists(fig_dir)) dir.create(fig_dir)

save_png <- function(plot, filename, width, height) {
  ggsave(
    file.path(fig_dir, filename),
    plot = plot, width = width, height = height,
    units = "in", dpi = 300, bg = PAPER
  )
}

# ---- THEME ----------------------------------------------------------------
# Sans-serif throughout. Minimal in-plot chrome — no titles, no subtitles,
# no captions inside the figure. Identification and explanation live in the
# LaTeX caption beneath each figure, in the same family.
theme_clean <- function(base_size = 12) {
  theme_minimal(base_size = base_size, base_family = "Helvetica") +
    theme(
      text = element_text(family = "Helvetica"),
      plot.title    = element_blank(),
      plot.subtitle = element_blank(),
      plot.caption  = element_blank(),
      plot.title.position = "plot",
      axis.title.x = element_text(size = base_size + 1, color = INK,
                                  margin = margin(t = 10)),
      axis.title.y = element_text(size = base_size + 1, color = INK,
                                  margin = margin(r = 10)),
      axis.text    = element_text(size = base_size, color = SLATE),
      axis.ticks   = element_line(color = SLATE, linewidth = 0.3),
      axis.ticks.length = unit(3, "pt"),
      panel.grid.minor   = element_blank(),
      panel.grid.major.x = element_line(color = GRID, linewidth = 0.3),
      panel.grid.major.y = element_blank(),
      panel.background = element_rect(fill = PAPER, color = NA),
      plot.background  = element_rect(fill = PAPER, color = NA),
      legend.position    = "bottom",
      legend.justification = "center",
      legend.title       = element_text(size = base_size, color = INK),
      legend.text        = element_text(size = base_size, color = INK),
      legend.key.height  = unit(12, "pt"),
      legend.key.width   = unit(16, "pt"),
      legend.margin      = margin(t = 12),
      legend.box.margin  = margin(t = 4),
      plot.margin = margin(t = 6, r = 16, b = 6, l = 6)
    )
}

# ---- HELPERS --------------------------------------------------------------
fmt_p_inline <- function(p) {
  if (p < 0.001) return("p < 0.001")
  paste0("p = ", format(round(p, 3), nsmall = 3))
}
```

```{r load_data}
# ============================================================================
# DATA — same cleaning as wcm_abstract_analysis_v4.Rmd
# >>> EDIT THIS PATH IF YOUR FILE LIVES SOMEWHERE ELSE <<<
# ============================================================================
data_path <- "~/Downloads/Tox Book.v2 (1).xlsx"

data <- read_excel(data_path, sheet = "INTOXICATE")
data <- data %>% mutate(row_id = row_number())

valid_categories <- c("Combination", "Analgesic", "Antidepressants",
                      "Street Drugs", "CO, As, CN", "Sedatives",
                      "Alcohol", "Unknown")

# Count "enrolled" as all rows in the registry that have any patient data
# (a row that is entirely NA is not really an enrolled patient — it's just
# an empty Excel row).  An enrolled patient has at minimum some identifying
# row content. This mirrors the abstract's "of 252 enrolled, 232 included"
# framing where the 20-patient drop was attributable to exclusion criteria.
n_enrolled <- data %>%
  filter(if_any(everything(), ~ !is.na(.))) %>%
  nrow()

clean_data <- data %>%
  filter(!is.na(Gender), !is.na(`Exposure Category`)) %>%
  mutate(
    Gender = trimws(Gender),
    `Exposure Category` = trimws(`Exposure Category`),
    `INTOXICATE SCORE`  = as.numeric(`INTOXICATE SCORE`),
    `Dichotomous Outcome` = case_when(
      tolower(`Actual Disposition`) == "icu" ~ 1L,
      TRUE ~ 0L
    ),
    Age = as.numeric(Age)
  ) %>%
  filter(Gender %in% c("M", "F"),
         `Exposure Category` %in% valid_categories,
         !is.na(Age), Age >= 12)

n_total    <- nrow(clean_data)
n_female   <- sum(clean_data$Gender == "F")
n_male     <- sum(clean_data$Gender == "M")
model_data <- clean_data %>% filter(!is.na(`INTOXICATE SCORE`))
n_model    <- nrow(model_data)

# Chi-squared on the 8x2 table for Figure 1 caption
cont_table <- table(clean_data$`Exposure Category`, clean_data$Gender)
chi_p <- suppressWarnings(chisq.test(cont_table))$p.value

# Sub-threshold male ICU rate for Figure 2 caption
male_icu_total <- model_data %>%
  filter(Gender == "M", `Dichotomous Outcome` == 1) %>%
  nrow()
male_icu_subthreshold <- model_data %>%
  filter(Gender == "M", `Dichotomous Outcome` == 1,
         `INTOXICATE SCORE` <= 6) %>%
  nrow()
```

\vspace{6pt}

\noindent\textbf{\large Figure 1. Exposure Category Counts by Biological Sex}

```{r fig1_paired_bars, fig.width=8.5, fig.height=5.2}
# ============================================================================
# FIGURE 1 — Paired horizontal bars (one row per sex per category)
# Structure mirrors the screenshot Yash referenced: each category gets two
# bars stacked vertically (F then M), categories ordered by total count.
# WCM palette substituted for the original black/gray.
# ============================================================================

fig1_data <- clean_data %>%
  count(`Exposure Category`, Gender) %>%
  group_by(`Exposure Category`) %>%
  mutate(total = sum(n),
         pct_female = sum(n[Gender == "F"]) / total * 100) %>%
  ungroup() %>%
  mutate(
    Sex = factor(Gender, levels = c("M", "F"),
                 labels = c("Male", "Female"))
  )

# Order categories by descending female percentage
cat_order <- fig1_data %>%
  distinct(`Exposure Category`, pct_female) %>%
  arrange(desc(pct_female)) %>%
  pull(`Exposure Category`)

fig1_data <- fig1_data %>%
  mutate(`Exposure Category` = factor(`Exposure Category`,
                                       levels = rev(cat_order)))

x_max <- max(fig1_data$n) * 1.10

fig1 <- ggplot(fig1_data,
               aes(x = n, y = `Exposure Category`, fill = Sex)) +
  geom_col(position = position_dodge(width = 0.78),
           width = 0.7, color = NA) +
  geom_text(aes(label = n),
            position = position_dodge(width = 0.78),
            hjust = -0.35, size = 3.6, color = INK,
            family = "Helvetica") +
  scale_fill_manual(values = c("Female" = WCM_MAROON,
                                "Male"   = WCM_NAVY),
                    breaks = c("Female", "Male"),
                    name = NULL) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.02)),
                     limits = c(0, x_max),
                     breaks = pretty(c(0, x_max))) +
  labs(x = "Number of patients", y = "Poisoning exposure category") +
  theme_clean(base_size = 12) +
  theme(
    axis.text.y = element_text(size = 12, color = INK,
                                margin = margin(r = 4))
  )

print(fig1)
save_png(fig1, "fig1_exposure_by_sex.png", width = 8.5, height = 5.2)
```

\begin{figure}[H]
\caption*{\textbf{Figure 1.} Number of female (maroon) and male (navy) patients in each poisoning exposure category. Categories ordered by descending female share. N = `r n_total` patients (`r n_female` female, `r n_male` male). Antidepressant exposures were predominantly female; street drug, CO/As/CN, and alcohol exposures were predominantly male. Chi-squared test of independence on the 8 \(\times\) 2 table: `r fmt_p_inline(chi_p)`.}
\end{figure}

\pagebreak

\noindent\textbf{\large Figure 2. INTOXICATE Score Distribution by Sex and ICU Outcome}

```{r fig2_score_distribution, fig.width=8.5, fig.height=4.8}
# ============================================================================
# FIGURE 2 — Jittered scatter, sex on y, INTOXICATE score on x, points
# colored by toxicologist disposition. Single dashed reference line at the
# original threshold (score > 6). No medians, no second threshold line, no
# in-plot title or annotations — keep it spare.
# ============================================================================

fig2_data <- model_data %>%
  mutate(
    Sex = factor(Gender, levels = c("F", "M"),
                 labels = c("Female", "Male")),
    Outcome = factor(`Dichotomous Outcome`, levels = c(1, 0),
                     labels = c("ICU", "Non-ICU"))
  )

# Group medians, computed per sex × outcome
group_medians <- fig2_data %>%
  group_by(Sex, Outcome) %>%
  summarise(med = median(`INTOXICATE SCORE`, na.rm = TRUE),
            .groups = "drop")

ICU_COLOR    <- "#E36C0A"   # orange
NONICU_COLOR <- "#7FB3D5"   # light steel blue

fig2 <- ggplot() +
  geom_vline(xintercept = 6, color = INK,
             linetype = "dashed", linewidth = 0.4) +
  # Patient points — circles
  geom_jitter(data = fig2_data,
              aes(x = `INTOXICATE SCORE`, y = Sex, color = Outcome,
                  shape = "Patient"),
              height = 0.22, alpha = 0.75, size = 2.4, stroke = 0) +
  # Group median triangles — nudged above/below the jitter band
  geom_point(data = group_medians,
             aes(x = med, y = Sex, color = Outcome,
                 shape = "Group median"),
             size = 5.5,
             position = position_nudge(
               y = ifelse(group_medians$Outcome == "ICU", 0.36, -0.36)
             )) +
  scale_color_manual(values = c("ICU" = ICU_COLOR,
                                "Non-ICU" = NONICU_COLOR),
                     name = "Toxicologist disposition") +
  scale_shape_manual(values = c("Patient" = 16, "Group median" = 17),
                     guide = "none") +
  scale_x_continuous(
    breaks = seq(0, max(fig2_data$`INTOXICATE SCORE`, na.rm = TRUE), 4),
    expand = expansion(add = c(0.5, 0.5))
  ) +
  scale_y_discrete(expand = expansion(add = c(0.7, 0.7))) +
  labs(x = "INTOXICATE score", y = NULL) +
  theme_clean(base_size = 12) +
  theme(
    axis.text.y = element_text(face = "bold", size = 13, color = INK)
  ) +
  guides(
    color = guide_legend(
      override.aes = list(shape = 16, size = 3.5, alpha = 1)
    )
  )

print(fig2)
save_png(fig2, "fig2_score_distribution.png", width = 8.5, height = 4.8)
```

\begin{figure}[H]
\caption*{\textbf{Figure 2.} INTOXICATE score for each patient with complete data (n = `r n_model`), stratified by biological sex and colored by toxicologist disposition. Triangles mark group medians for ICU and non-ICU dispositions within each sex. The dashed line marks the original threshold of IRS \(>\) 6. `r male_icu_subthreshold` of `r male_icu_total` male patients (`r round(100*male_icu_subthreshold/male_icu_total)`\%) with raw scores at or below the threshold ultimately required ICU-level care, motivating a sex-specific adjustment to the score.}
\end{figure}

\pagebreak

\noindent\textbf{\large Cohort statistics for poster Results section}

```{r results_stats, results='asis'}
# ============================================================================
# COHORT STATS — every number needed to populate the Results section
# of the poster, formatted to mirror the abstract's prose style.
# ============================================================================

library(stats)

# --- Sensitivity & specificity, with and without sex (+3 male adjustment) ---
wilson_ci <- function(x, n) {
  if (n == 0) return(c(0, 0, 0))
  z <- 1.96; p <- x / n
  denom  <- 1 + z^2 / n
  center <- (p + z^2 / (2 * n)) / denom
  spread <- z * sqrt((p * (1 - p) + z^2 / (4 * n)) / n) / denom
  c(p, max(0, center - spread), min(1, center + spread))
}

scores  <- model_data$`INTOXICATE SCORE`
genders <- model_data$Gender
actual  <- model_data$`Dichotomous Outcome`

pred_base <- ifelse(scores > 6, 1, 0)
sens_base <- wilson_ci(sum(pred_base == 1 & actual == 1), sum(actual == 1))
spec_base <- wilson_ci(sum(pred_base == 0 & actual == 0), sum(actual == 0))

adj_scores <- ifelse(genders == "M", scores + 3, scores)
pred_sex   <- ifelse(adj_scores > 6, 1, 0)
sens_sex   <- wilson_ci(sum(pred_sex == 1 & actual == 1), sum(actual == 1))
spec_sex   <- wilson_ci(sum(pred_sex == 0 & actual == 0), sum(actual == 0))

# --- OR for male sex ---
glm_fit <- glm(`Dichotomous Outcome` ~ `INTOXICATE SCORE` + Gender,
               data = model_data, family = binomial)
glm_ci   <- suppressMessages(confint(glm_fit))
or_male  <- exp(coef(glm_fit)["GenderM"])
or_lo    <- exp(glm_ci["GenderM", 1])
or_hi    <- exp(glm_ci["GenderM", 2])

# --- Median age + IQR by sex ---
age_by_sex <- clean_data %>%
  group_by(Gender) %>%
  summarise(
    n   = n(),
    med = median(Age, na.rm = TRUE),
    q1  = quantile(Age, 0.25, na.rm = TRUE),
    q3  = quantile(Age, 0.75, na.rm = TRUE),
    .groups = "drop"
  )
m_row <- age_by_sex %>% filter(Gender == "M")
f_row <- age_by_sex %>% filter(Gender == "F")

# --- Variance inflation factors for collinearity check ---
# Build a design matrix matching the full IRS regression (sex + IRS score
# is the minimal model; we extend to a richer set if those columns exist).
vif_max <- NA_real_
tryCatch({
  if (requireNamespace("car", quietly = TRUE)) {
    vif_fit <- glm(`Dichotomous Outcome` ~ `INTOXICATE SCORE` + Gender,
                   data = model_data, family = binomial)
    vif_vals <- car::vif(vif_fit)
    vif_max <- max(vif_vals)
  }
}, error = function(e) NULL)

# --- Exposure-by-sex percentages for the three categories the abstract
#     highlights (so we can drop them straight into the prose) ---
exp_pcts <- clean_data %>%
  count(`Exposure Category`, Gender) %>%
  group_by(`Exposure Category`) %>%
  mutate(pct = round(100 * n / sum(n))) %>%
  ungroup()

get_pct <- function(cat, sex) {
  v <- exp_pcts %>% filter(`Exposure Category` == cat, Gender == sex) %>%
    pull(pct)
  if (length(v) == 0) NA_integer_ else v
}

ad_f   <- get_pct("Antidepressants", "F")
ad_m   <- get_pct("Antidepressants", "M")
sd_m   <- get_pct("Street Drugs",    "M")
co_m   <- get_pct("CO, As, CN",      "M")

# --- Print everything ---
cat("\\textbf{Drop-in numbers for Results section:}\\\\\n\n")

cat(sprintf(
  "Of %d enrolled patients, %d were included (%d male, median age %.0f [IQR %.0f--%.0f]; %d female, median age %.0f [IQR %.0f--%.0f]).\\\\\n\n",
  n_enrolled, n_total,
  m_row$n, m_row$med, m_row$q1, m_row$q3,
  f_row$n, f_row$med, f_row$q1, f_row$q3
))

cat(sprintf(
  "Including biological sex increased sensitivity from %.0f\\%% (95\\%% CI %.0f--%.0f\\%%) to %.0f\\%% (%.0f--%.0f\\%%) and decreased specificity from %.0f\\%% (%.0f--%.0f\\%%) to %.0f\\%% (%.0f--%.0f\\%%).\\\\\n\n",
  100 * sens_base[1], 100 * sens_base[2], 100 * sens_base[3],
  100 * sens_sex[1],  100 * sens_sex[2],  100 * sens_sex[3],
  100 * spec_base[1], 100 * spec_base[2], 100 * spec_base[3],
  100 * spec_sex[1],  100 * spec_sex[2],  100 * spec_sex[3]
))

cat(sprintf(
  "Male sex was associated with increased odds of ICU-level care (OR %.2f, 95\\%% CI %.2f--%.2f); biological sex was associated with exposure category (chi-squared %s)",
  or_male, or_lo, or_hi, fmt_p_inline(chi_p)
))
if (!is.na(vif_max)) {
  cat(sprintf(
    " but not strongly collinear with model predictors (VIF $<$ %.0f).\\\\\n\n",
    ceiling(vif_max)
  ))
} else {
  cat(". (Note: install the \\texttt{car} package to compute VIF.)\\\\\n\n")
}

if (!is.na(ad_f) && !is.na(sd_m) && !is.na(co_m)) {
  cat(sprintf(
    "Antidepressant exposures were more common in females (%d\\%% vs %d\\%%). Street drugs (%d\\%%) and CO/As/CN exposures (%d\\%%) were more common in males.\\\\\n\n",
    ad_f, ad_m, sd_m, co_m
  ))
}

cat("\\vspace{8pt}\n\n")
cat("\\textbf{Sanity-check counts:}\\\\\n")
cat(sprintf("Enrolled (non-empty rows): %d\\\\\n", n_enrolled))
cat(sprintf("Included after exclusions: %d\\\\\n", n_total))
cat(sprintf("Excluded: %d\\\\\n", n_enrolled - n_total))
cat(sprintf("With complete IRS data (used for model): %d\\\\\n", n_model))
```

\begin{quote}
\noindent\textit{Note: \texttt{n\_enrolled} is computed as the number of non-empty rows in the INTOXICATE sheet. If your registry's true enrollment count (before any data entry) lives elsewhere, override the printed sentence with the correct figure.}
\end{quote}
