---
title: "INTOXICATE Poster Figures"
author: "Yash Jaggi"
date: "WCM EM Research Symposium · May 7, 2026"
output:
  pdf_document:
    latex_engine: xelatex
    toc: false
geometry: "left=0.4in,right=0.4in,top=0.6in,bottom=0.6in"
header-includes:
  - \usepackage{float}
  - \usepackage{caption}
  - \captionsetup[figure]{labelformat=empty,font=normalsize,skip=4pt}
  - \renewcommand{\familydefault}{\sfdefault}
  - \pagestyle{empty}
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(readxl)
library(scales)
library(stringr)
library(knitr)
library(kableExtra)

knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  fig.align = "center", dpi = 300
)

# ---- PALETTE ---------------------------------------------------------------
WCM_MAROON   <- "#A8132B"
FEMALE_COLOR <- "#E36C0A"
MALE_COLOR   <- "#7FB3D5"
INK          <- "#1A1A1A"
SLATE        <- "#4A4A4A"
PAPER        <- "#FFFFFF"
GRID         <- "#E8E8E8"

fig_dir <- "poster_figures"
if (!dir.exists(fig_dir)) dir.create(fig_dir)

save_png <- function(plot, filename, width, height) {
  ggsave(file.path(fig_dir, filename), plot = plot,
         width = width, height = height, units = "in", dpi = 300, bg = PAPER)
}

theme_clean <- function(base_size = 12) {
  theme_minimal(base_size = base_size, base_family = "Helvetica") +
    theme(
      plot.title    = element_blank(),
      plot.subtitle = element_blank(),
      plot.caption  = element_blank(),
      axis.title.x  = element_text(size = base_size + 1, color = INK,
                                   margin = margin(t = 10)),
      axis.title.y  = element_text(size = base_size + 1, color = INK,
                                   margin = margin(r = 10)),
      axis.text     = element_text(size = base_size, color = SLATE),
      axis.ticks    = element_line(color = SLATE, linewidth = 0.3),
      panel.grid.minor   = element_blank(),
      panel.grid.major.x = element_line(color = GRID, linewidth = 0.3),
      panel.grid.major.y = element_blank(),
      panel.background   = element_rect(fill = PAPER, color = NA),
      plot.background    = element_rect(fill = PAPER, color = NA),
      legend.position    = "bottom",
      legend.title       = element_text(size = base_size, color = INK),
      legend.text        = element_text(size = base_size, color = INK),
      plot.margin = margin(t = 6, r = 16, b = 6, l = 6)
    )
}

fmt_p_inline <- function(p) {
  if (is.na(p)) return("NA")
  if (p < 0.001) return("p < 0.001")
  paste0("p = ", format(round(p, 3), nsmall = 3))
}

# Wilson score CI — returns named vector c(est, lo, hi)
wilson_ci <- function(x, n, conf = 0.95) {
  if (n == 0) return(c(est = 0, lo = 0, hi = 0))
  z <- qnorm(1 - (1 - conf) / 2)
  p <- x / n
  denom  <- 1 + z^2 / n
  center <- (p + z^2 / (2 * n)) / denom
  spread <- z * sqrt((p * (1 - p) + z^2 / (4 * n)) / n) / denom
  c(est = p,
    lo  = max(0, center - spread),
    hi  = min(1, center + spread))
}

# Cohen's kappa — always returns a plain list, never errors
compute_kappa <- function(pred, actual) {
  # Point estimate via contingency table (no external package needed)
  tab <- table(factor(pred, levels = c(0, 1)),
               factor(actual, levels = c(0, 1)))
  n  <- sum(tab)
  po <- sum(diag(tab)) / n
  pe <- sum(rowSums(tab) * colSums(tab)) / n^2
  k  <- if (pe == 1) NA_real_ else (po - pe) / (1 - pe)

  # CI via psych if installed, otherwise NA
  lo <- NA_real_; hi <- NA_real_
  if (requireNamespace("psych", quietly = TRUE) && !is.na(k)) {
    tryCatch({
      kk <- psych::cohen.kappa(cbind(pred, actual))
      lo <- kk$confid["unweighted kappa", "lower"]
      hi <- kk$confid["unweighted kappa", "upper"]
    }, error = function(e) NULL)
  }
  list(kappa = k, lo = lo, hi = hi)
}

# Format a metric with optional CI
fmt_metric <- function(est, lo, hi, pct = FALSE) {
  scale <- if (pct) 100 else 1
  fmt   <- if (pct) "%.0f%%" else "%.2f"
  if (is.na(lo) || is.na(hi)) {
    sprintf(fmt, est * scale)
  } else {
    sprintf(paste0(fmt, " (", fmt, "\u2013", fmt, ")"),
            est * scale, lo * scale, hi * scale)
  }
}
```

```{r load_data}
# >>> EDIT PATH IF NEEDED <<<
data_path <- "~/Downloads/Tox Book.v2 (1).xlsx"

data <- read_excel(data_path, sheet = "INTOXICATE")
data <- data %>% mutate(row_id = row_number())

valid_categories <- c("Combination", "Analgesic", "Antidepressants",
                      "Street Drugs", "CO, As, CN", "Sedatives",
                      "Alcohol", "Unknown")

n_enrolled <- data %>%
  filter(if_any(everything(), ~ !is.na(.))) %>%
  nrow()

clean_data <- data %>%
  filter(!is.na(Gender), !is.na(`Exposure Category`)) %>%
  mutate(
    Gender                = trimws(Gender),
    `Exposure Category`   = trimws(`Exposure Category`),
    `INTOXICATE SCORE`    = as.numeric(`INTOXICATE SCORE`),
    `Dichotomous Outcome` = case_when(
      tolower(`Actual Disposition`) == "icu" ~ 1L, TRUE ~ 0L),
    Age = as.numeric(Age),
    SBP = as.numeric(SBP)          # fix: force SBP to numeric
  ) %>%
  filter(Gender %in% c("M", "F"),
         `Exposure Category` %in% valid_categories,
         !is.na(Age), Age >= 12)

n_total  <- nrow(clean_data)
n_female <- sum(clean_data$Gender == "F")
n_male   <- sum(clean_data$Gender == "M")

# Model data: complete cases across all regression variables
model_data <- clean_data %>%
  mutate(
    Pulse               = as.numeric(Pulse),
    GCS                 = as.numeric(GCS),
    `Exposure Category` = factor(`Exposure Category`),
    Gender              = factor(Gender, levels = c("F", "M"))
  ) %>%
  filter(
    !is.na(`INTOXICATE SCORE`),
    !is.na(Pulse), !is.na(SBP), !is.na(GCS),
    !is.na(`Respiratory Insufficiency`),
    !is.na(Cirrhosis),
    !is.na(`Secondary Reason for ICU Admission`),
    !is.na(Dysrhythmia),
    !is.na(`Exposure Category`)
  )

n_model <- nrow(model_data)

cont_table <- table(clean_data$`Exposure Category`, clean_data$Gender)
chi_p <- suppressWarnings(chisq.test(cont_table))$p.value
```

\pagebreak

\noindent\textbf{\large Model Comparison: IRS With vs. Without Biological Sex}

\noindent\textit{Two logistic regression models fit on the same complete-case data. All predictors match the supervisor-specified model. Biological sex is added directly as a covariate in the second model.}

\vspace{6pt}

```{r model_fit, results='asis'}
model_wo_sex <- glm(
  `Dichotomous Outcome` ~ Pulse + SBP + GCS + Age +
    `Respiratory Insufficiency` + Cirrhosis +
    `Secondary Reason for ICU Admission` + Dysrhythmia +
    `Exposure Category`,
  data = model_data, family = binomial
)

model_w_sex <- glm(
  `Dichotomous Outcome` ~ Gender + Pulse + SBP + GCS + Age +
    `Respiratory Insufficiency` + Cirrhosis +
    `Secondary Reason for ICU Admission` + Dysrhythmia +
    `Exposure Category`,
  data = model_data, family = binomial
)

# Helper: clean coefficient table from a glm object
coef_table <- function(fit) {
  s <- summary(fit)$coefficients
  df <- data.frame(
    Variable    = rownames(s),
    Coefficient = round(s[, "Estimate"], 4),
    SE          = round(s[, "Std. Error"], 4),
    `z-value`   = round(s[, "z value"], 4),
    `p-value`   = ifelse(s[, "Pr(>|z|)"] < 0.001, "< 0.001",
                          sprintf("%.4f", s[, "Pr(>|z|)"])),
    check.names = FALSE
  )
  rownames(df) <- NULL
  df
}

cat("\\textbf{Model without sex:}\\vspace{4pt}\n\n")
kbl(coef_table(model_wo_sex), booktabs = TRUE,
    align = c("l","r","r","r","r")) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"),
                full_width = FALSE, position = "left") %>%
  print()
cat("\n\n")

cat("\\textbf{Model with sex:}\\vspace{4pt}\n\n")
kbl(coef_table(model_w_sex), booktabs = TRUE,
    align = c("l","r","r","r","r")) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"),
                full_width = FALSE, position = "left") %>%
  print()
cat("\n\n")

# Likelihood ratio test
lrt    <- anova(model_wo_sex, model_w_sex, test = "LRT")
lrt_p  <- lrt[2, "Pr(>Chi)"]
lrt_x2 <- lrt[2, "Deviance"]
lrt_df <- lrt[2, "Df"]

lrt_df_out <- data.frame(
  `Chi-squared` = round(lrt_x2, 3),
  df            = lrt_df,
  `p-value`     = ifelse(lrt_p < 0.001, "< 0.001",
                          sprintf("%.3f", lrt_p)),
  Significant   = ifelse(lrt_p < 0.05, "Yes", "No"),
  check.names   = FALSE
)

cat("\\textbf{Likelihood ratio test (with sex vs. without):}\\vspace{4pt}\n\n")
kbl(lrt_df_out, booktabs = TRUE, align = c("r","r","r","c")) %>%
  kable_styling(latex_options = "hold_position",
                full_width = FALSE, position = "left") %>%
  print()
cat("\n\n")
```

\pagebreak

\noindent\textbf{\large Performance Metrics: Sensitivity, Specificity, and Cohen's Kappa}

\vspace{6pt}

```{r performance_metrics, results='asis'}
actual <- model_data$`Dichotomous Outcome`
n_pos  <- sum(actual == 1)
n_neg  <- sum(actual == 0)

pred_wo_prob <- predict(model_wo_sex, type = "response")
pred_w_prob  <- predict(model_w_sex,  type = "response")
pred_wo      <- ifelse(pred_wo_prob >= 0.5, 1L, 0L)
pred_w       <- ifelse(pred_w_prob  >= 0.5, 1L, 0L)

# Confusion matrix counts
tp_wo <- sum(pred_wo == 1 & actual == 1)
fp_wo <- sum(pred_wo == 1 & actual == 0)
tn_wo <- sum(pred_wo == 0 & actual == 0)
fn_wo <- sum(pred_wo == 0 & actual == 1)

tp_w  <- sum(pred_w  == 1 & actual == 1)
fp_w  <- sum(pred_w  == 1 & actual == 0)
tn_w  <- sum(pred_w  == 0 & actual == 0)
fn_w  <- sum(pred_w  == 0 & actual == 1)

# Metrics with CIs
sens_wo <- wilson_ci(tp_wo, tp_wo + fn_wo)
spec_wo <- wilson_ci(tn_wo, tn_wo + fp_wo)
kap_wo  <- compute_kappa(pred_wo, actual)

sens_w  <- wilson_ci(tp_w,  tp_w  + fn_w)
spec_w  <- wilson_ci(tn_w,  tn_w  + fp_w)
kap_w   <- compute_kappa(pred_w, actual)

# --- Performance summary table (poster table) --------------------------------
perf <- data.frame(
  Metric = c("Sensitivity", "Specificity", "Cohen's Kappa"),
  `Excluding Biological Sex` = c(
    fmt_metric(sens_wo["est"], sens_wo["lo"], sens_wo["hi"]),
    fmt_metric(spec_wo["est"], spec_wo["lo"], spec_wo["hi"]),
    fmt_metric(kap_wo$kappa,  kap_wo$lo,     kap_wo$hi)
  ),
  `Including Biological Sex` = c(
    fmt_metric(sens_w["est"], sens_w["lo"], sens_w["hi"]),
    fmt_metric(spec_w["est"], spec_w["lo"], spec_w["hi"]),
    fmt_metric(kap_w$kappa,  kap_w$lo,     kap_w$hi)
  ),
  check.names = FALSE
)

cat("\\textbf{Performance summary (threshold = 0.5, 95\\% CI):}\\vspace{4pt}\n\n")
kbl(perf, booktabs = TRUE, align = c("l","c","c")) %>%
  kable_styling(latex_options = "hold_position",
                full_width = FALSE, position = "left") %>%
  print()
cat("\n\n")

# --- Confusion matrices (two separate clean tables) -------------------------
cat("\\textbf{Confusion matrix — excluding biological sex:}\\vspace{4pt}\n\n")
conf_wo <- data.frame(
  ` `              = c("Predicted non-ICU", "Predicted ICU"),
  `Actual non-ICU` = c(tn_wo, fp_wo),
  `Actual ICU`     = c(fn_wo, tp_wo),
  check.names = FALSE
)
kbl(conf_wo, booktabs = TRUE, align = c("l","r","r")) %>%
  kable_styling(latex_options = "hold_position",
                full_width = FALSE, position = "left") %>%
  print()
cat("\n\n")

cat("\\textbf{Confusion matrix — including biological sex:}\\vspace{4pt}\n\n")
conf_w <- data.frame(
  ` `              = c("Predicted non-ICU", "Predicted ICU"),
  `Actual non-ICU` = c(tn_w, fp_w),
  `Actual ICU`     = c(fn_w, tp_w),
  check.names = FALSE
)
kbl(conf_w, booktabs = TRUE, align = c("l","r","r")) %>%
  kable_styling(latex_options = "hold_position",
                full_width = FALSE, position = "left") %>%
  print()
cat("\n\n")

# --- Per-sex breakdown -------------------------------------------------------
cat("\\textbf{Per-sex breakdown (both models):}\\vspace{4pt}\n\n")

sex_perf <- model_data %>%
  mutate(
    pred_wo = ifelse(pred_wo_prob >= 0.5, 1L, 0L),
    pred_w  = ifelse(pred_w_prob  >= 0.5, 1L, 0L)
  ) %>%
  group_by(Gender) %>%
  summarise(
    n          = n(),
    actual_icu = sum(`Dichotomous Outcome` == 1),
    tp_wo = sum(pred_wo == 1 & `Dichotomous Outcome` == 1),
    fp_wo = sum(pred_wo == 1 & `Dichotomous Outcome` == 0),
    tn_wo = sum(pred_wo == 0 & `Dichotomous Outcome` == 0),
    fn_wo = sum(pred_wo == 0 & `Dichotomous Outcome` == 1),
    tp_w  = sum(pred_w  == 1 & `Dichotomous Outcome` == 1),
    fp_w  = sum(pred_w  == 1 & `Dichotomous Outcome` == 0),
    tn_w  = sum(pred_w  == 0 & `Dichotomous Outcome` == 0),
    fn_w  = sum(pred_w  == 0 & `Dichotomous Outcome` == 1),
    .groups = "drop"
  ) %>%
  mutate(
    Sex              = ifelse(Gender == "F", "Female", "Male"),
    `Sens w/o sex`   = sprintf("%.0f%% (%d/%d)",
                                100 * tp_wo / (tp_wo + fn_wo),
                                tp_wo, tp_wo + fn_wo),
    `Spec w/o sex`   = sprintf("%.0f%% (%d/%d)",
                                100 * tn_wo / (tn_wo + fp_wo),
                                tn_wo, tn_wo + fp_wo),
    `Over-triage w/o`= sprintf("%d (%.0f%%)", fp_wo, 100 * fp_wo / n),
    `Sens w/ sex`    = sprintf("%.0f%% (%d/%d)",
                                100 * tp_w / (tp_w + fn_w),
                                tp_w, tp_w + fn_w),
    `Spec w/ sex`    = sprintf("%.0f%% (%d/%d)",
                                100 * tn_w / (tn_w + fp_w),
                                tn_w, tn_w + fp_w),
    `Over-triage w/`  = sprintf("%d (%.0f%%)", fp_w, 100 * fp_w / n)
  ) %>%
  select(Sex, n, `Actual ICU` = actual_icu,
         `Sens w/o sex`, `Spec w/o sex`, `Over-triage w/o`,
         `Sens w/ sex`,  `Spec w/ sex`,  `Over-triage w/`)

kbl(sex_perf, booktabs = TRUE,
    align = c("l","r","r","r","r","r","r","r","r")) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"),
                full_width = FALSE, position = "left") %>%
  add_header_above(c(" " = 3,
                     "Excl. Biological Sex" = 3,
                     "Incl. Biological Sex" = 3)) %>%
  print()
cat("\n\n")

# Headline delta
delta_k <- kap_w$kappa - kap_wo$kappa
cat(sprintf(
  "\\textbf{Key result:} Cohen's kappa changed from %.2f to %.2f (delta = %+.2f). LRT: $\\chi^2$ = %.2f, df = %d, %s.\n\n",
  kap_wo$kappa, kap_w$kappa, delta_k, lrt_x2, lrt_df, fmt_p_inline(lrt_p)
))
```

\pagebreak

\noindent\textbf{\large Cohort Statistics for Poster Results Section}

```{r results_stats, results='asis'}
# Use model_data (n=187 complete cases) for all reported medians
age_by_sex <- model_data %>%
  group_by(Gender) %>%
  summarise(n   = n(),
            med = median(Age, na.rm = TRUE),
            q1  = quantile(Age, 0.25, na.rm = TRUE),
            q3  = quantile(Age, 0.75, na.rm = TRUE),
            .groups = "drop")
m_row <- age_by_sex %>% filter(Gender == "M")
f_row <- age_by_sex %>% filter(Gender == "F")
n_model_m <- m_row$n
n_model_f <- f_row$n

# OR for male sex from model with sex
sex_coef <- coef(summary(model_w_sex))["GenderM", ]
or_male  <- exp(sex_coef["Estimate"])
or_ci    <- suppressMessages(confint(model_w_sex))["GenderM", ]
or_lo    <- exp(or_ci[1]); or_hi <- exp(or_ci[2])
sex_p    <- sex_coef["Pr(>|z|)"]

# Exposure percentages from model_data cohort
exp_pcts <- model_data %>%
  count(`Exposure Category`, Gender) %>%
  group_by(`Exposure Category`) %>%
  mutate(pct = round(100 * n / sum(n))) %>%
  ungroup()

get_pct <- function(cat, sex) {
  v <- exp_pcts %>%
    filter(`Exposure Category` == cat, Gender == sex) %>%
    pull(pct)
  if (length(v) == 0) NA_integer_ else v
}

ad_f <- get_pct("Antidepressants", "F"); ad_m <- get_pct("Antidepressants", "M")
sd_m <- get_pct("Street Drugs", "M");    co_m <- get_pct("CO, As, CN", "M")

# Male sensitivity values for bullet point
male_sens_wo <- sex_perf %>% filter(Sex == "Male") %>% pull(`Sens w/o sex`)
male_sens_w  <- sex_perf %>% filter(Sex == "Male") %>% pull(`Sens w/ sex`)

cat("\\textbf{Cohort and drop-in bullets for Results:}\\vspace{4pt}\n\n")

# Cohort line
cat(sprintf(
  "\\textbf{Cohort line:} %d patients with complete model data (%d male, median age %.0f [IQR %.0f--%.0f]; %d female, median age %.0f [IQR %.0f--%.0f]).\n\n",
  n_model,
  m_row$n, m_row$med, m_row$q1, m_row$q3,
  f_row$n, f_row$med, f_row$q1, f_row$q3
))

# Bullet points
cat("\\textbf{Bullet points:}\n\n")
cat(sprintf(
  "\\textbullet\\ Including biological sex significantly improved overall model agreement with toxicologist-recommended disposition (Cohen's kappa %.2f to %.2f, LRT %s).\n\n",
  kap_wo$kappa, kap_w$kappa, fmt_p_inline(lrt_p)
))
cat(sprintf(
  "\\textbullet\\ Sensitivity for identifying ICU-level patients increased from %.0f\\%% to %.0f\\%%, driven by improved detection in the male cohort (%s to %s).\n\n",
  100 * sens_wo["est"], 100 * sens_w["est"],
  male_sens_wo, male_sens_w
))
cat(sprintf(
  "\\textbullet\\ Male sex was independently associated with %.1f-fold increased odds of ICU-level care (OR %.2f, 95\\%% CI %.2f--%.2f, %s).\n\n",
  or_male, or_male, or_lo, or_hi, fmt_p_inline(sex_p)
))
cat(sprintf(
  "\\textbullet\\ Biological sex was significantly associated with exposure category (%s): antidepressant exposures were more common in females (%d\\%%), while street drug and CO/As/CN exposures were more common in males (%d\\%% and %d\\%% respectively).\n\n",
  fmt_p_inline(chi_p), ad_f, sd_m, co_m
))

cat("\\vspace{6pt}\n\n")
cat("\\textbf{Sanity-check counts:}\n\n")
cat(sprintf("Enrolled (non-empty rows): %d\\\\\n", n_enrolled))
cat(sprintf("Included after exclusions: %d\\\\\n", n_total))
cat(sprintf("Complete model data (used for regression): %d\\\\\n", n_model))
```

\pagebreak

\noindent\textbf{\large Figure 1. IRS Performance With vs. Without Biological Sex}

```{r fig1_metrics, fig.width=7.0, fig.height=4.5}
metrics_df <- data.frame(
  Metric = factor(
    rep(c("Sensitivity", "Specificity", "Cohen's Kappa"), 2),
    levels = c("Sensitivity", "Specificity", "Cohen's Kappa")
  ),
  Model = factor(
    rep(c("Excluding Biological Sex", "Including Biological Sex"), each = 3),
    levels = c("Excluding Biological Sex", "Including Biological Sex")
  ),
  est = c(sens_wo["est"], spec_wo["est"], kap_wo$kappa,
          sens_w["est"],  spec_w["est"],  kap_w$kappa),
  lo  = c(sens_wo["lo"],  spec_wo["lo"],
          ifelse(is.na(kap_wo$lo), kap_wo$kappa - 0.05, kap_wo$lo),
          sens_w["lo"],   spec_w["lo"],
          ifelse(is.na(kap_w$lo),  kap_w$kappa  - 0.05, kap_w$lo)),
  hi  = c(sens_wo["hi"],  spec_wo["hi"],
          ifelse(is.na(kap_wo$hi), kap_wo$kappa + 0.05, kap_wo$hi),
          sens_w["hi"],   spec_w["hi"],
          ifelse(is.na(kap_w$hi),  kap_w$kappa  + 0.05, kap_w$hi))
)

fig1 <- ggplot(metrics_df, aes(x = Metric, y = est, fill = Model)) +
  geom_col(position = position_dodge(width = 0.65), width = 0.60,
           color = NA) +
  geom_errorbar(aes(ymin = lo, ymax = hi),
                position = position_dodge(width = 0.65),
                width = 0.18, linewidth = 0.7, color = INK) +
  geom_text(aes(label = sprintf("%.2f", est), y = hi + 0.03),
            position = position_dodge(width = 0.65),
            size = 3.5, color = INK) +
  scale_fill_manual(
    values = c("Excluding Biological Sex" = MALE_COLOR,
               "Including Biological Sex" = FEMALE_COLOR),
    name = NULL
  ) +
  scale_y_continuous(limits = c(0, 1.08),
                     breaks = seq(0, 1, 0.2),
                     labels = scales::number_format(accuracy = 0.1),
                     expand = expansion(mult = c(0, 0))) +
  labs(title = "Figure 1. Biological Sex Improves IRS Prediction of ICU Disposition",
       x = NULL, y = NULL) +
  theme_clean(base_size = 12) +
  theme(
    plot.title         = element_text(face = "bold", size = 13, color = INK,
                                      margin = margin(b = 10)),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = GRID, linewidth = 0.4),
    legend.position    = "bottom"
  )

print(fig1)
save_png(fig1, "fig1_model_comparison.png", width = 7.0, height = 4.5)
```

\begin{figure}[H]
\caption*{\textbf{Figure 1.} Sensitivity, specificity, and Cohen's kappa for IRS logistic regression models excluding (blue) and including (orange) biological sex as a direct covariate, compared against toxicologist-recommended disposition. Error bars show 95\% confidence intervals (Wilson score for sensitivity/specificity; psych package for kappa). Threshold = 0.5 on predicted probability. N = `r n_model` patients with complete model data.}
\end{figure}

\pagebreak

\noindent\textbf{\large Figure 2. Exposure Category Counts by Biological Sex}

```{r fig2_exposure, fig.width=7.5, fig.height=5.5}
fig2_data <- clean_data %>%
  count(`Exposure Category`, Gender) %>%
  group_by(`Exposure Category`) %>%
  mutate(total = sum(n)) %>%
  ungroup() %>%
  mutate(Sex = factor(Gender, levels = c("F", "M"),
                      labels = c("Female", "Male")))

cat_order <- fig2_data %>%
  distinct(`Exposure Category`, total) %>%
  arrange(desc(total)) %>%
  pull(`Exposure Category`)

fig2_data <- fig2_data %>%
  mutate(`Exposure Category` = factor(`Exposure Category`,
                                       levels = rev(cat_order)))

x_max <- max(fig2_data$n) * 1.10
chi_label <- sprintf("Sex \u00d7 Exposure Category: %s", fmt_p_inline(chi_p))

fig2 <- ggplot(fig2_data,
               aes(x = n, y = `Exposure Category`, fill = Sex)) +
  geom_col(position = position_dodge(width = 0.78), width = 0.7, color = NA) +
  geom_text(aes(label = n),
            position = position_dodge(width = 0.78),
            hjust = -0.35, size = 3.6, color = INK) +
  annotate("label",
           x = x_max * 0.98, y = length(cat_order) / 2 + 0.5,
           label = chi_label,
           hjust = 1, size = 3.8, color = INK,
           label.size = 0.3, label.r = unit(2, "pt"), fill = "#F5F5F5") +
  scale_fill_manual(values = c("Female" = FEMALE_COLOR, "Male" = MALE_COLOR),
                    name = NULL) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.02)),
                     limits = c(0, x_max),
                     breaks = pretty(c(0, x_max))) +
  scale_y_discrete(expand = expansion(add = c(0.6, 2.2))) +
  labs(title = "Figure 2. Exposure Category Counts by Biological Sex",
       x = "Number of Patients", y = NULL) +
  theme_clean(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold", size = 13, color = INK,
                                   margin = margin(b = 10)),
    legend.position = c(0.82, 0.92),
    legend.background = element_rect(fill = "white", color = NA),
    axis.text.y     = element_text(size = 12, color = INK)
  )

print(fig2)
save_png(fig2, "fig2_exposure_by_sex.png", width = 8.5, height = 5.5)
```

\begin{figure}[H]
\caption*{\textbf{Figure 2.} Number of female (orange) and male (blue) patients in each poisoning exposure category, ordered by descending total count. N = `r n_total` patients (`r n_female` female, `r n_male` male). Chi-squared test: `r fmt_p_inline(chi_p)`.}
\end{figure}
