---
title: "Supporting Material for NACCT 2024 Abstract"
output : html_document
fontsize: 12pt
---

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(dplyr)
library(yardstick)
library(gt)
library(irr)
library(verification)
library(ggplot2)
library(gtsummary)
library(cutpointr)

df  <-  read_excel("../data/snapshot.05062024.xlsx", sheet = "INTOXICATE", range = cell_cols("A:O"))

xf <-  df %>%
  mutate(`Actual Disposition`  = case_when(
    `Actual Disposition` == "ICU" ~ "ICU",
    `Actual Disposition` == "GMF" ~ "Not ICU",
    `Actual Disposition` == "Discharge" ~ "Not ICU",
    TRUE ~ "NA"
  ), 
  `Predicted Disposition`  = case_when(
    `Predicted Disposition` == "ICU" ~ "ICU",
    `Predicted Disposition` == "GMF" ~ "Not ICU",
    `Predicted Disposition` == "Discharge" ~ "Not ICU",
    TRUE ~ "NA"))

xtab  <- xf %>%
 tabyl(`Actual Disposition`,`Predicted Disposition`)
```
```{r table, echo=FALSE, fig.cap="INTOXICATE Predicted vs Actual Disposition", fig.height=4, fig.width=6}


xf %>%
  filter(`Predicted Disposition` %in% c("ICU","Not ICU")) %>%
  tbl_cross(`Actual Disposition`,`Predicted Disposition`, percent="col") %>%
  bold_labels()

xf %>%
  tabyl(`Actual Disposition`,`Predicted Disposition`) %>%
  adorn_totals("row") %>%
  adorn_totals("col") %>%
  gt() %>%
  cols_move("NA", after = "Not ICU") %>%
  tab_header(title = "INTOXICATE vs Clinician Judgement") %>%
  tab_spanner(label = "INTOXICATE", 
    columns=c("ICU","Not ICU","NA")) %>%
  tab_footnote(html("ICU, Intensive Care Unit; Not ICU, Discharged,<br>Transferred to Psychiatry, or Admitted to<br>a General Medical Floor")) %>%
  tab_spanner(label = "Predicted Disposition")
```
<br>

```{r combined table, echo=FALSE}

act_icu = df %>%
  drop_na("INTOXICATE SCORE") %>%
  filter(`Actual Disposition`=="ICU") %>%
  nrow()

act_gmf = df %>%
  drop_na("INTOXICATE SCORE") %>%
  filter(`Actual Disposition`=="GMF") %>%
  nrow()
   
 
```
```

`````

*Results.*  We included `r nrow(df)` patients, aged `r quantile(df$Age, c(0.50))` [`r quantile(df$Age, c(0.25))` to `r quantile(df$Age, c(0.75))`] years, `r round(100*sum(df$Gender=="M")/nrow(df))`% male, `r round(100*sum(df$Gender=="F")/nrow(df))`% female. We excluded `r sum(is.na(df$"Predicted Disposition"))` patients because the chart did not contain all information needed to predict disposition or the the patient's poisoning was not a main contributor to disposition.  Of the remaining `r df %>% drop_na("INTOXICATE SCORE") %>% nrow()` patients, `r act_icu` were admitted to the ICU,
`r act_gmf` to the floor, and `r sum(df$"Actual Disposition"=="Discharge")` discharged. INTOXICATE predicted that `r sum(df$"Predicted Disposition"=="ICU" & df$"Actual Disposition"=="GMF", na.rm=TRUE)` of `r sum(df$"Actual Disposition"=="GMF", na.rm=TRUE)` patients admitted to the floor should be admitted to the ICU and that `r sum(df$"Predicted Disposition"=="ICU" & df$"Actual Disposition"=="Discharge", na.rm=TRUE)` of the `r sum(df$"Predicted Disposition"=="GMF", na.rm=TRUE)` patients discharged should be admitted to the ICU. None of these were transferred from the floor to the ICU or Stepdown Unit. Cohen\'s
kappa was `r round(kappa2(xf[,c("Predicted Disposition","Actual Disposition")])$value,4)`. All records contained the fields needed to calculate the
INTOXICATE risk score and inter-rater reliability, demonstrating no agreement above chance between INTOXICATE's recommendations and those of bedside toxicologists.


#### ROC Curve

```{r ROC, echo=FALSE, fig.cap="INTOXICATE ROC Curve", fig.height=4, fig.width=6}

 xf %>%
    mutate(`Actual Disposition` = as.factor(`Actual Disposition`)) %>%
    roc_curve(`Actual Disposition`, `INTOXICATE SCORE`) %>%
    ggplot(aes(x=1-specificity, y=sensitivity)) +
    geom_path() +
    geom_abline(lty=3) + 
    coord_equal() +
    theme_bw()
```

`````
#### Cutpoint Analysis

```{r cutpoint, }
cp  <-  xf %>%
       cutpointr(`INTOXICATE SCORE`, `Actual Disposition`, method = maximize_metric, metric = sum_sens_spec, pos_class = "ICU", neg_class="Not ICU", na.rm=TRUE)

summary(cp)
```


