---
title: "Recalibration of INTOXICATE Model"
output : 
  html_document:
    keep_md: true
fontsize: 12pt
---

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(dplyr)
library(yardstick)
library(gt)
library(ggrepel)
library(ggplot2)
library(ggpubr)
library(gtsummary)
library(psych)
library(caret)

# Make sure you're using rms, not a masked function
detach("package:psych", unload=TRUE)
library(rms)
library(ggplotify)
library(rmda)
library(pROC)
library(cowplot)
knitr::opts_chunk$set(dev = "png",
                      dpi = 300,
                      echo = FALSE,
                      cache = TRUE)
df  <-  read_excel("../data/snapshots/snapshot.04192025.intoxicate.xlsx", sheet = "INTOXICATE", range = cell_cols("A:AC"))

df <- df %>%
  mutate(age_category = case_when(
    Age < 12 ~ "Pediatric",
    Age >= 12 & Age < 18 ~ "Adolescent",
    Age >= 18 ~ "Adult"
  ))

xf <-  df %>%
  mutate(`Actual Disposition`  = case_when(
    `Actual Disposition` == "ICU" ~ "ICU",
    `Actual Disposition` == "GMF" ~ "Not ICU",
    `Actual Disposition` == "Discharge" ~ "Not ICU",
    TRUE ~ "NA"
  ), 
  `Predicted Disposition`  = case_when(
    `Predicted Disposition` == "ICU" ~ "ICU",
    `Predicted Disposition` == "GMF" ~ "Not ICU",
    `Predicted Disposition` == "Discharge" ~ "Not ICU",
    TRUE ~ "NA"))


xf$icu_admit <- ifelse(xf$`Actual Disposition` == "ICU", 1, 0)
dd <- rms::datadist(xf[ , sapply(xf, function(v) length(na.omit(unique(v))) > 1)])
options(datadist = "dd")
xf$pred_admit <- ifelse(xf$`Predicted Disposition` == "ICU", 1, 0)
```

```{r recalibration}

cal_model <- lrm(`Actual Disposition` ~ `INTOXICATE SCORE`, data = xf, x = TRUE, y = TRUE)

# Calibration plot
cal <- calibrate(cal_model, method = "boot", B = 200)
plot(cal, 
     xlab = "Predicted Probability", 
     ylab = "Observed Probability",
     main = "Calibration Plot for INTOXICATE in ED Population")


# Decision Curve Analysis
dca_result <- decision_curve(
  icu_admit ~ `INTOXICATE SCORE`,
  data = xf,
  family = binomial(link = "logit"),
  thresholds = seq(0, 1, by = 0.01),
  confidence.intervals = 0.95,
  study.design = "cohort"
)

# Plot decision curves
plot_decision_curve(
  dca_result,
  curve.names = "INTOXICATE Score",
  xlab = "Threshold Probability for ICU Admission",
  ylab = "Net Benefit",
  legend.position = "topright"
)

# Recalibration model: refit outcome vs. original predicted log-odds
recal_model <- glm(
  icu_admit ~ `INTOXICATE SCORE`,
  data = xf %>% filter(Age > 12),
  family = binomial(link = "logit")
)

summary(recal_model)

ggplot(xf, aes(x = `INTOXICATE SCORE`, y = icu_admit)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"),
              se = TRUE, colour = "red", fill = "pink", alpha = 0.4) +
  geom_line(aes(y = pred_admit), colour = "blue") +
  labs(title = "Recalibration of INTOXICATE Score in ED Population",
       x = "INTOXICATE Risk Score",
       y = "Predicted Probability of ICU Admission") +
  theme_classic()

xf$ed_prob <- plogis(-2.96 + 0.138 * xf$`INTOXICATE SCORE`)
xf$ed_pred_disp <- ifelse(xf$ed_prob > 0.16, "ICU", "Not ICU")
xf %>% 
  filter(Age > 12) %>%
  tbl_cross(`Predicted Disposition`, `Actual Disposition`)

xf %>% 
  filter(Age > 12) %>%
  tbl_cross(ed_pred_disp, `Actual Disposition`)

# --- Compute confusion matrices safely ---
cm_original <- xf %>%
  filter(Age > 12) %>%
  { caret::confusionMatrix(
      as.factor(.$`Predicted Disposition`),
      as.factor(.$`Actual Disposition`),
      positive = "ICU"
    )
  }

cm_recal <- xf %>%
  filter(Age > 12) %>%
  { caret::confusionMatrix(
      as.factor(.$ed_pred_disp),
      as.factor(.$`Actual Disposition`),
      positive = "ICU"
    )
  }

cm_adj <- xf %>%
  filter(Age > 12) %>%
  mutate(
    adj_pred_disp = ifelse(`INTOXICATE SCORE` > 17, "ICU", "Not ICU")
  ) %>%
  { caret::confusionMatrix(
      as.factor(.$adj_pred_disp),
      as.factor(.$`Actual Disposition`),
      positive = "ICU"
    )
  }

# --- Convert to tidy data frames ---
tidy_cm <- function(cm, label) {
  as.data.frame(cm$byClass) |>
    tibble::rownames_to_column("Metric") |>
    dplyr::mutate(Model = label)
}

df_cm <- bind_rows(
  tidy_cm(cm_original, "Original INTOXICATE"),
  tidy_cm(cm_adj, "Adjusted Threshold (17)"),
  tidy_cm(cm_recal, "Recalibrated (ED)")
)

df_cm %>%
  pivot_wider(names_from = Metric, values_from = "cm$byClass") %>%
  select(Model, Sensitivity, Specificity, `Balanced Accuracy`, `Pos Pred Value`, `Neg Pred Value`) %>%
  gt() 
