---
title: "Recalibration of INTOXICATE Model"
output : 
  html_document:
    keep_md: true
fontsize: 12pt
---

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(dplyr)
library(yardstick)
library(gt)
library(ggrepel)
library(ggplot2)
library(ggpubr)
library(gtsummary)
library(stringr)
library(psych)
library(caret)

# Make sure you're using rms, not a masked function
detach("package:psych", unload=TRUE)
library(rms)
library(ggplotify)
library(rmda)
library(pROC)
library(cowplot)
knitr::opts_chunk$set(dev = "png",
                      dpi = 300,
                      echo = FALSE,
                      cache = FALSE)

#was 04192025
df  <-  read_excel("../data/snapshots/snapshot.04302026.xlsx", sheet = "INTOXICATE", range = cell_cols("A:AC"))

df <- df %>%
  mutate(age_category = case_when(
    Age < 12 ~ "Pediatric",
    Age >= 12 & Age < 18 ~ "Adolescent",
    Age >= 18 ~ "Adult"
  ))


xf <- df %>%
   mutate(`INTOXICATE SCORE` = as.numeric(`INTOXICATE SCORE`)) %>%
  filter(Age > 12, !is.na(Age), Gender %in% c("M","F"), !is.na(`Actual Disposition`), !is.na(`INTOXICATE SCORE`)) %>%
  mutate(
    `Exposure Category` = str_squish(`Exposure Category`),
    `Exposure Category` = case_when(
      `Exposure Category` %in% c(
        "clorox bleach",
        "Chlorine Gas"
      ) ~ "Not Classified",

      `Exposure Category` %in% c(
        "Sedative (Combination)",
        "Polysubstance"
      ) ~ "Combination",

      `Exposure Category` == "Antipsychotic" ~ "Antidepressants",

      TRUE ~ `Exposure Category`
    ),
  `Actual Disposition`  = case_when(
    `Actual Disposition` == "ICU" ~ "ICU",
    `Actual Disposition` == "GMF" ~ "Not ICU",
    `Actual Disposition` == "No ICU" ~ "Not ICU",
    str_detect(`Actual Disposition`, "(?i)Discharge") ~ "Not ICU",
    str_detect(`Actual Disposition`, "(?i)Psych") ~ "Not ICU",
    str_detect(`Actual Disposition`, "(?i)Admit to Neuro") ~ "Not ICU"),
    `Predicted Disposition` = case_when(
      `INTOXICATE SCORE` > 6 ~ "ICU",
      `INTOXICATE SCORE` <= 6 ~ "Not ICU",
      TRUE ~ NA_character_),)
xf$icu_admit <- ifelse(xf$`Actual Disposition` == "ICU", 1, 0)
dd <- rms::datadist(xf[ , sapply(xf, function(v) length(na.omit(unique(v))) > 1)])
options(datadist = "dd")
xf$pred_admit <- ifelse(xf$`Predicted Disposition` == "ICU", 1, 0)
```

```{r tbl1}

xf %>%
  select(Age, Gender, `Exposure Category`, `INTOXICATE SCORE`, `Actual Disposition`,`Predicted Disposition`) %>%
  tbl_summary(by = Gender, missing = "no") %>%
  add_overall() %>%
  add_p() %>%
  add_n()

```

```{r recalibration}

cal_model <- lrm(`Actual Disposition` ~ `INTOXICATE SCORE`, data = xf, x = TRUE, y = TRUE)

# Calibration plot
cal <- calibrate(cal_model, method = "boot", B = 200)
plot(cal, 
     xlab = "Predicted Probability", 
     ylab = "Observed Probability",
     main = "Calibration Plot for INTOXICATE in ED Population")


# Decision Curve Analysis
dca_result <- decision_curve(
  icu_admit ~ `INTOXICATE SCORE`,
  data = xf,
  family = binomial(link = "logit"),
  thresholds = seq(0, 1, by = 0.01),
  confidence.intervals = 0.95,
  study.design = "cohort"
)

# Plot decision curves
plot_decision_curve(
  dca_result,
  curve.names = "INTOXICATE Score",
  xlab = "Threshold Probability for ICU Admission",
  ylab = "Net Benefit",
  legend.position = "topright"
)

# Recalibration model: refit outcome vs. original predicted log-odds
recal_model <- glm(
  icu_admit ~ `INTOXICATE SCORE`,
  data = xf %>% filter(Age > 12),
  family = binomial(link = "logit")
)

summary(recal_model)

ggplot(xf, aes(x = `INTOXICATE SCORE`, y = icu_admit)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"),
              se = TRUE, colour = "red", fill = "pink", alpha = 0.4) +
  geom_line(aes(y = pred_admit), colour = "blue") +
  labs(title = "Recalibration of INTOXICATE Score in ED Population",
       x = "INTOXICATE Risk Score",
       y = "Predicted Probability of ICU Admission") +
  theme_classic()

xf$ed_prob <- plogis(-2.96 + 0.138 * xf$`INTOXICATE SCORE`)
xf$ed_pred_disp <- ifelse(xf$ed_prob > 0.16, "ICU", "Not ICU")
xf %>% 
  filter(Age > 12) %>%
  tbl_cross(`Predicted Disposition`, `Actual Disposition`) %>%
  modify_header(label = "**Original INTOXICATE**") 

xf %>% 
  filter(Age > 12) %>%
  tbl_cross(ed_pred_disp, `Actual Disposition`) %>%
  modify_header(label = "**Recalibrated ED**") 

# --- Predictions ---

xf <- xf %>%
  mutate(
    adj_pred_disp = ifelse(`INTOXICATE SCORE` > 17, "ICU", "Not ICU"),
    adj_pred_disp_sex = ifelse(
      `INTOXICATE SCORE` - 0.5 * (Gender == "M") > 17,
      "ICU",
      "Not ICU"
    ),
    ed_prob = predict(recal_model, newdata = xf, type = "response"),
    ed_pred_disp = ifelse(ed_prob > 0.16, "ICU", "Not ICU"),
    ed_prob_sex = predict(recal_model_sex_exposure, newdata = xf, type = "response"),
    ed_pred_disp_sex = ifelse(ed_prob_sex > 0.16, "ICU", "Not ICU"),
    orig_prob_sex = plogis(`INTOXICATE SCORE` - 0.5 * (Gender == "M")),
    orig_pred_disp_sex = ifelse(orig_prob_sex > 0.16, "ICU", "Not ICU")
  )

# Create confusion matrices and calculate metrics for each model

xf %>%
  filter(Age > 12) %>%
  tbl_cross(adj_pred_disp, `Actual Disposition`) %>%
  add_overall() %>%
    modify_header(label = "**Adjusted Threshold (>17)**")

xf %>%
  filter(Age > 12) %>%
  tbl_cross(adj_pred_disp_sex, `Actual Disposition`) %>%
  add_overall() %>%
    modify_header(label = "**Adjusted Threshold (>17) + Gender**") 

xf %>%
  filter(Age > 12) %>%
  tbl_cross(ed_pred_disp_sex, `Actual Disposition`) %>%
  add_overall() %>%
    modify_header(label = "**Recalibrated ED + Gender**")

# --- Helper functions ---

boot_metrics <- function(data, pred_col, ref_col, B = 1000) {

  n <- nrow(data)

  boot_stats <- replicate(B, {
    idx <- sample(seq_len(n), replace = TRUE)
    d <- data[idx, ]

    cm <- caret::confusionMatrix(
      factor(d[[pred_col]], levels = c("Not ICU", "ICU")),
      factor(d[[ref_col]], levels = c("Not ICU", "ICU")),
      positive = "ICU"
    )

    tab <- cm$table

    TP <- tab["ICU", "ICU"]
    FN <- tab["Not ICU", "ICU"]
    TN <- tab["Not ICU", "Not ICU"]
    FP <- tab["ICU", "Not ICU"]

    sens <- ifelse((TP + FN) > 0, TP / (TP + FN), NA)
    spec <- ifelse((TN + FP) > 0, TN / (TN + FP), NA)
    ppv  <- ifelse((TP + FP) > 0, TP / (TP + FP), NA)
    npv  <- ifelse((TN + FN) > 0, TN / (TN + FN), NA)

    kappa <- as.numeric(cm$overall["Kappa"])

    c(sens, spec, ppv, npv, kappa)
  })

  boot_stats <- t(boot_stats)

  colnames(boot_stats) <- c("Sensitivity", "Specificity", "PPV", "NPV", "Kappa")

  point_est <- colMeans(boot_stats, na.rm = TRUE)

  ci <- apply(boot_stats, 2, quantile, probs = c(0.025, 0.975), na.rm = TRUE)

  list(
    est = point_est,
    ci = ci,
    boot = boot_stats
  )
}

make_row <- function(data, pred_col, label) {

  out <- boot_metrics(data, pred_col, "Actual Disposition")

  tibble(
    Model = label,
    Sensitivity = out$est["Sensitivity"],
    Sensitivity_CI = sprintf("%.2f–%.2f", out$ci[1, "Sensitivity"], out$ci[2, "Sensitivity"]),
    Specificity = out$est["Specificity"],
    Specificity_CI = sprintf("%.2f–%.2f", out$ci[1, "Specificity"], out$ci[2, "Specificity"]),
    PPV = out$est["PPV"],
    PPV_CI = sprintf("%.2f–%.2f", out$ci[1, "PPV"], out$ci[2, "PPV"]),
    NPV = out$est["NPV"],
    NPV_CI = sprintf("%.2f–%.2f", out$ci[1, "NPV"], out$ci[2, "NPV"]),
    Kappa = out$est["Kappa"],
    Kappa_CI = sprintf("%.2f–%.2f", out$ci[1, "Kappa"], out$ci[2, "Kappa"]),
    boot = list(out$boot)
  )
}

make_cm <- function(pred, ref) {
  caret::confusionMatrix(
    factor(pred, levels = c("Not ICU", "ICU")),
    factor(ref, levels = c("Not ICU", "ICU")),
    positive = "ICU"
  )
}

get_metrics_ci <- function(data, pred_col, label) {
  cm <- make_cm(data[[pred_col]], data$`Actual Disposition`)
  tab <- cm$table

  TP <- tab["ICU", "ICU"]
  FN <- tab["Not ICU", "ICU"]
  TN <- tab["Not ICU", "Not ICU"]
  FP <- tab["ICU", "Not ICU"]

  sens <- TP / (TP + FN)
  spec <- TN / (TN + FP)

  sens_ci <- binom.test(TP, TP + FN)$conf.int
  spec_ci <- binom.test(TN, TN + FP)$conf.int
  ppv_ci  <- binom.test(TP, TP + FP)$conf.int
  npv_ci  <- binom.test(TN, TN + FN)$conf.int

  # --- Cohen’s kappa ---
  kappa <- as.numeric(cm$overall["Kappa"])

  # SE for Fleiss kappa
  n <- sum(tab)
  po <- sum(diag(tab)) / n
  pe <- sum(rowSums(tab) * colSums(tab)) / n^2
  se_kappa <- sqrt((po * (1 - po)) / (n * (1 - pe)^2))

  kappa_ci <- c(
    kappa - 1.96 * se_kappa,
    kappa + 1.96 * se_kappa
  )

  tibble(
    Model = label,
    N = n,
    Sensitivity = sens,
    Sensitivity_CI = sprintf("%.2f–%.2f", sens_ci[1], sens_ci[2]),
    Specificity = spec,
    Specificity_CI = sprintf("%.2f–%.2f", spec_ci[1], spec_ci[2]),
    `Balanced Accuracy` = (sens + spec) / 2,
    PPV = TP / (TP + FP),
    PPV_CI = sprintf("%.2f–%.2f", ppv_ci[1], ppv_ci[2]),
    NPV = TN / (TN + FN),
    NPV_CI = sprintf("%.2f–%.2f", npv_ci[1], npv_ci[2]),
    Kappa = kappa,
    Kappa_CI = sprintf("%.2f–%.2f", kappa_ci[1], kappa_ci[2])
  )
}
# --- Single performance table ---

df_perf <- bind_rows(
  get_metrics_ci(xf, "Predicted Disposition", "Original INTOXICATE"),
  get_metrics_ci(xf, "orig_pred_disp_sex", "Original INTOXICATE + Gender"),
  get_metrics_ci(xf, "adj_pred_disp", "Adjusted Threshold (>17)"),
  get_metrics_ci(xf, "adj_pred_disp_sex", "Adjusted Threshold (>17) + Gender"),
  get_metrics_ci(xf, "ed_pred_disp", "Recalibrated ED"),
  get_metrics_ci(xf, "ed_pred_disp_sex", "Recalibrated ED + Gender")
) %>%
  mutate(
    Comparator = case_when(
      Model == "Original INTOXICATE + Gender" ~ "Original INTOXICATE",
      Model == "Adjusted Threshold (>17) + Gender" ~ "Adjusted Threshold (>17)",
      Model == "Recalibrated ED + Gender" ~ "Recalibrated ED",
      TRUE ~ NA_character_
    )
  ) %>%
  left_join(
    select(., Comparator = Model, Comparator_Kappa = Kappa),
    by = "Comparator"
  ) %>%
  mutate(
    Delta_Kappa = Kappa - Comparator_Kappa
  )

rows <- list(
  make_row(xf, "Predicted Disposition", "Original"),
  make_row(xf, "orig_pred_disp_sex", "Original + Gender"),
  make_row(xf, "adj_pred_disp", "Adjusted"),
  make_row(xf, "adj_pred_disp_sex", "Adjusted + Gender"),
  make_row(xf, "ed_pred_disp", "Recalibrated"),
  make_row(xf, "ed_pred_disp_sex", "Recalibrated + Gender")
)

df_perf <- bind_rows(rows)

compute_delta_kappa <- function(row1, row2) {
  # row1 = base, row2 = +gender
  delta <- row2$boot[[1]][, "Kappa"] - row1$boot[[1]][, "Kappa"]

  c(
    mean = mean(delta, na.rm = TRUE),
    lower = quantile(delta, 0.025, na.rm = TRUE),
    upper = quantile(delta, 0.975, na.rm = TRUE)
  )
}

# Map comparisons
delta_list <- list(
  compute_delta_kappa(rows[[1]], rows[[2]]), # Original
  compute_delta_kappa(rows[[3]], rows[[4]]), # Adjusted
  compute_delta_kappa(rows[[5]], rows[[6]])  # Recalibrated
)

delta_df <- tibble(
  Model = c(
    "Original + Gender",
    "Adjusted + Gender",
    "Recalibrated + Gender"
  ),
  Delta_Kappa = sapply(delta_list, function(x) x["mean"]),
  Delta_Kappa_CI = sapply(delta_list, function(x)
    sprintf("%.2f–%.2f", x["lower"], x["upper"])
  )
)

df_perf <- df_perf %>%
  left_join(delta_df, by = "Model")

df_perf %>%
  mutate(
    Sensitivity = sprintf("%.2f (%s)", Sensitivity, Sensitivity_CI),
    Specificity = sprintf("%.2f (%s)", Specificity, Specificity_CI),
    PPV = sprintf("%.2f (%s)", PPV, PPV_CI),
    NPV = sprintf("%.2f (%s)", NPV, NPV_CI),
    Kappa = sprintf("%.2f (%s)", Kappa, Kappa_CI),
    `Δ Kappa vs no gender` = ifelse(
      is.na(Delta_Kappa),
      "—",
      sprintf("%+.2f", Delta_Kappa)
    ),
  ) %>%
  select(
    Model,
    Sensitivity,
    Specificity,
    PPV,
    NPV,
    Kappa,
    `Δ Kappa vs no gender`
  ) %>%
  gt()
